Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-03-17 paper.tex

\textbf{Evaluation without Caching.}
Since our first version of @aeval@ implements the nondeterministic
evaluation, given the defined @nd@, the evaluation function @aeval@ can be easily
defined. In the case that @e@ is a @let@ expression,
we invoke @nd@ for twice and perform a nesting two-step, depth-first evaluation
over possible values of @App(f, ae)@.

@@

In this section, we present the second version that uses delimited controls.

@@

To address this problem, we introduce two operators @nd@ and @ndcps@ for handling non-deterministic choices.

\textbf{Nondeterminism Operators, Again.}
The first operator @nd@ is simplified from the previous one, but specialized with type @Ans@ and @Cache@. 
We may still keep the types to be generic, but the purpose of presentation clearly, exposing @Cache@ explictly
reveals that cache should be monotonically accumulated to the rest of computation.
Now @nd@ only takes a function @f@ that applies on type @(T, Cache)@ where value of type @T@ comes from the set,
and @nd@ does recursive call on itself over elements in the set @ts@.
We also move the accumulation operation (@acc++f(ts.head)@) inside of @nd@.

@@

The second operator @ndcps@ is implemented with delimited controls but simply calls our 
first operator @nd@ in body.
\begin{lstlisting}
def ndcps[T](ts: Set[T], acc: Ans): (T, Cache) @cps[Ans] = shift { f: (((T, Cache)) => Ans) =>
  nd(ts, acc, f)
}
\end{lstlisting}
The function @ndcps@ takes two arguments: a set of elements with type @T@, 
and an initial accumulated value of type @Ans@.
The @ndcps@ iteratively returns an element from the set which the element has type @T@ and annotated by
\texttt{\@ cps[Ans]}. This annotation denotes the final returned value of @ndcps@ is type @Ans@.
The @f@ introduced by the @shift@ operator is the delimited continuation from the call-site of @ndcps@.

\textbf{}
We add an annotation \texttt{ cps[Ans]} on the return type of the @aeval@ function.
Now, for every nondeterministic choice in the abstract interpreter, we can simply call 
@ndcps@ on them and write the program sequentially as a concrete interpreter.
@ndcps@ also returns the latest cache to its left-hand side definition.
The cache is still accumulated to the next call to @ndcps@ or @aeval@, 
given the fact that the abstract interpreter should always use the latest cache.

@@

The function @analyze@ is also changed by adding a @reset@ around the function call to @aeval@.
The afterwards updating on cache becomes sequential.

@@

Our work establish the pushdown control-flow analysis through refunctionalization and a
proper caching algorithm. The call/return is naturally matched in the stack of 
the meta language.

>>>>

\textbf{Evaluation without Caching.}
Since our first version of @aeval@ implements the nondeterministic
evaluation, given the defined @nd@, the evaluation function @aeval@ can be easily
defined. In the case that @e@ is a @let@ expression,
we invoke @nd@ for twice and perform a nesting two-step, depth-first evaluation
over possible values of @App(f, ae)@.
The values of branches from on fork point will be accumulated, and then
returned to the upper fork point.

@@

In this section, we present the second version that uses delimited control operators.

@@

To address this problem, we reintroduce two operators @nd@ and @ndcps@ for handling non-deterministic choices
in a way that mimics extended continuation-passing style.

\textbf{Nondeterminism Operators, Again.}
The first operator @nd@ is simplified from the previous one, but specialized with type @Ans@ and @Cache@. 
We may still keep the types to be generic, but the purpose of presentation clearly, exposing @Cache@ explictly
reveals that cache should be monotonically accumulated to the rest of computation.
Now @nd@ only takes a function @f@ that applies on type @(T, Cache)@ where value of type @T@ comes from the set,
and @nd@ does recursive call on itself over elements in the set @ts@.
We also move the accumulation operation (@++@) inside of @nd@.

@@

The second operator @ndcps@ is implemented with delimited control operator @shift@ but simply calls our 
first operator @nd@ in body.
\begin{lstlisting}
def ndcps[T](ts: Set[T], acc: Ans): (T, Cache) @cps[Ans] = shift { f: (((T, Cache)) => Ans) =>
  nd(ts, acc, f)
}
\end{lstlisting}
The function @ndcps@ takes two arguments: a set of elements with type @T@, 
and an initial accumulated value of type @Ans@.
The @ndcps@ iteratively returns an element from the set along with the latest cache 
, and the return type is also annotated by \verb|@cps[Ans]|. 
This annotation denotes the final returned value of @ndcps@ is type @Ans@.
The @f@ introduced by the @shift@ operator is the delimited continuation from the call-site of @ndcps@.

Now we can use these two operators to redefine @aeval@.
We add an annotation \verb|@cps[Ans]| on the return type of the @aeval@ function.
Then, for every nondeterministic choice in the abstract interpreter, we can simply call 
@ndcps@ on the set of possible choices and write the program sequentially as a concrete interpreter.
@ndcps@ also returns the latest cache to its left-hand side definition.
The cache is still accumulated to the next call to @ndcps@ or @aeval@, 
given the fact that the abstract interpreter should always use the latest cache.

@@

The function @analyze@ is also changed by adding a @reset@ operator around the function call to @aeval@.
The afterwards updating on cache becomes sequential.

@@

Our work establish the pushdown control-flow analysis through refunctionalization and a
proper caching algorithm. The call/return is naturally matched by the stack of 
the meta language.
